Stage-Splitting a Modal Language

### 2014-05-24 paper/sigplanconf-template.tex
Added figures for grammar, values, types, statics, and dynamics.
Added a section about PE->SS.  Also a splitting figure.

\usepackage{amsmath}
\usepackage{stmaryrd}

@@

@@

We can recover a valid partial evaluator from a stage-splitter by stage-splitting the input function $f$ into $f_1$ and $f_2$, computing $\llbracket f_1 \rrbracket(x)$ to obtain $\bar x$, and then returning an $f_x$ such that $\llbracket f_x \rrbracket(y) = \llbracket f_2 \rrbracket(\bar x, y)$.  However, one {\em cannot} construct a valid stage-splitter out of a partial evaluator, since a partial evaluator by definition depends on the first-stage input, whereas a stage-splitter does not have that available.

Note that this does not mean that stage-splitting is a strict generalization of partial evaluation.  In practice, partial evaluators easily perform optimizations (such as branch elimination, discussed later) which are beyond the scope of stage-splitting, and would require further technology than has been developed here.  It is best to think of stage-splitting as simply the first half of partial evaluation, where the back half is an optimizer. [Might be able to come up with a futamura projection-like statement here, which would be really really really cool.]

@@

Present the splitting judgement.  Give statements of type and value correctness for splitting.  Give all of the splitting rules.  Talk through a few of them.

>>>>

\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{bbm}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{mathpartir}
\usepackage{mathabx}

@@

\input {macros}
\input {grammars}
\input {semantics}
\input {splitting}

@@

\subsection{Partial Evaluator from Stage-Splitter}

We can recover a valid partial evaluator from a stage-splitter by stage-splitting the input function $f$ into $f_1$ and $f_2$, computing $\llbracket f_1 \rrbracket(x)$ to obtain $\bar x$, and then returning an $f_x$ such that $\llbracket f_x \rrbracket(y) = \llbracket f_2 \rrbracket(\bar x, y)$.  Note that this does not mean that stage-splitting is a strict generalization of partial evaluation.  In practice, partial evaluators easily perform optimizations (such as branch elimination, discussed later) which are beyond the scope of stage-splitting, and would require further technology than has been developed here.  It is best to think of stage-splitting as simply the first half of partial evaluation, where the back half is an optimizer. [Might be able to come up with a futamura projection-like statement here, which would be really really really cool.]

\subsection{Stage-Splitter from Partial Evaluator}

Likewise, we can easily recover a stage-splitter from a partial evaluator.  If $p$ is a valid partial evaluator, then we can define a stage splitter $s$ such that $s(f)=(f_1,f_2)$, where
\begin{align*}
[f_1](x) &= p (f,x) \\
[f_2](l,y) &= [l] (y)
\end{align*}
This implicitly requires that the languages in which $f_1$ and $f_2$ are expressed are strong enough to write a partial evaluator, but that is the case in this paper.  A stage-splitter defined this way leaves much to be desired.  Firstly, partial evaluation of $f$ may be too expensive for the context in which $f_1$ needs to run.  Additionally, the intermediate data structure created this way may be much larger than necessary, as it would contain all of the residual code.

@@

[Present the splitting judgement.  Give statements of type and value correctness for splitting.  Give all of the splitting rules.  Talk through a few of them.]

The theorem that we want to be true is...

\begin{center}
\begin{tabular}{l}
If $\splittwo{e}{A}{p,l.r}$ \\
and $\reducetwo{e}{v}$ \\
then $\reduce{p}{u}$ \\
and $\reduce{[u/l]r}{w}$ \\
and $v = w$ 
\end{tabular}
\end{center}

%\begin{center}
%\begin{tabular}{l}
%If $\splitone {e}{A}{c,l.r}$ \\
%and $\colone {\gamma_1}{\Gamma}$ \\
%and $\coltwo {\gamma_2}{\Gamma}$\\
%and $\rsone{\gamma_1(e)}{v}{q}$ \\
%and $\reducetwo{\gamma_2(q)}{w}$ \\
%then $\reduce{|\gamma_1|(c)}{\valprod{v}{u}}$ \\
%and $\reduce{[u/l]|\gamma_2|(r)}{w}$
%\end{tabular}
%~~~
%\begin{tabular}{l}
%If $\splittwo{e}{B}{p,l.r}$ \\
%and $\colone {\gamma_1}{\Gamma}$ \\
%and $\coltwo {\gamma_2}{\Gamma}$\\
%and $\rstwo{\gamma_1(e)}{q}$ \\
%and $\reducetwo{\gamma_2(q)}{w}$ \\
%then $\reduce{|\gamma_1|(p)}{u}$ \\
%and $\reduce{[u/l]|\gamma_2|(r)}{w}$ 
%\end{tabular}
%\end{center}
